Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-06 measurements.tex
Corrections to format writing

The authors implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the Raspberry Pi 2 Model B,
which features four ARM Cortex-A7 cores in a Broadcom BCM2836 \ac{SOC}
with a 900 MHz clock. Each core has a 32 KiB L1 data cache and 32 KiB
L1 instruction cache. The cores share a 512 KiB L2 cache. All the
measured results, including the baseline results, were obtained with
NEON-enabled code adopted from the Fifi library [\textbf{Fifi reference}].
The NEON extension provides an 128-bit \ac{SIMD} instruction set to the
\ac{Raspi} 2. Figures~\ref{enc_dec1024}, \ref{enc_dec128} and
\ref{enc_dec16} show the encoding and decoding throughput in MiB per
second for different generation sizes ($g$ = 1024, 128, and 16
respectively). The throughput is the rate of generating $g$ coded packets,
divided by the time the encoder took to perform the task. The size of
each coded packet was fixed to $1536$~bytes since that is the typical
size of an Ethernet frame. The blocked operations were performed dividing
the matrices in squared sub-blocks of $16,\ 32\ ,\ 64,\ldots,\ 1024$ operands
(words in the finite field) in height and width. The figures show only
block sizes of $16 \times 16$, $32 \times 32$, $64 \times 64$ operands
since with bigger block sizes, the operands do not fit in L1 cache.

The following test cases were considered:

\subsubsection{Baseline encoding}
The baseline results involve no recording of the
\ac{DAG} and are performed in a \emph{by-the-book} fashion. The encoder uses
only one thread. And the difference between the non-blocked and blocked
encoding schemes is that in the blocked scheme, the matrix multiplications are
performed dividing the matrices in sub-blocks in order to make the algorithm
cache efficient as described in Section~\ref{sub:implementation-multicore}.

\subsubsection{Encoding blocked}
The encoding results were performed using the
method described in Section~\ref{sub:implementation-multicore}. The time
recorded includes the dependencies resolving, creation of the \ac{DAG},
and the task scheduling. In practice, it would suffice to calculate and
store this information only once per generation size.

@@

For $g= 1024$, the blocked baseline measurements outperforms the non blocked
variant. This means that making the matrix multiplication algorithm cache
efficient brings an increase in throughput by a factor of 2.83. When using the
algorithm described in Section~\ref{sub:implementation-multicore},
encoding with four cores is on average 3.7 times faster than with one core.
Similarly, decoding with four codes is 3.05 times faster, on average, than
decoding with a single core. Figure~\ref{enc_dec1024} shows that the
implemented algorithm, by exploiting cache efficiency and only three extra
cores provides a gain of 10 folds compared with traditional non-blocked
algorithms. With $g = 1024$, the matrix inversion becomes more expensive
than at smaller generations sizes. Therefore, the decoding throughput
is 46\% of the encoding throughput.

@@

For $g$ = 128, the differences between the baselines operations show that a
blocked algorithm is 15\% faster than the non-blocked variant. Encoding with
four cores is 2.82 times faster than with a single core. Due to the smaller
matrix sizes, the gain when using blocked operations in the baselines is not
that significant when compared with $g = 1024$. For the same reason, the matrix
inversion is less expensive. As a consequence, the decoding throughput is 71\%
of the encoding throughput.

@@

The differences in throughput, for all generation sizes, between the
blocked baseline and the single threaded scheduled measurements are due the
time spent resolving the dependencies and the scheduling overhead. These
effects are negligible for big generation sizes, while considerable for
small matrices.

@@

To compare how much slower is the matrix multiplication with respect to the
matrix inversion for different generation sizes, we ran a set of tests. We used
a single core to perform the operations. We changed the generation sizes,
performed matrices multiplications and matrix inversions, and measured the time
spent doing so which we name $T_{mult}$ and $T_{inv}$. We calculate the ratio
between these two measured times defined as $r = \frac{T_{mult}}{T_{inv}}$.
Table~\ref{runtimes} summarizes the results. The bigger the matrix size, the
smaller is the calculated ratio. This means that when the problems are bigger,
the decoding throughput decreases compared with the encoding throughput.

>>>>

The authors implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the \ac{Raspi} 2 model,
which features four ARMv7 cores in a Broadcom BCM2836 \ac{SOC}
with a 900 MHz clock. Each core has a 32 KiB L1 data cache and 32 KiB
L1 instruction cache. The cores share a 512 KiB L2 cache. All the
measured results, including the baseline results, were obtained with
NEON-enabled code adapted from the Fifi library [\textbf{Fifi reference}].
The \ac{Raspi} 2 has a NEON extension instruction set which provides
an 128-bit \ac{SIMD} instructions that speed the computations.
Figures~\ref{enc_dec1024}, \ref{enc_dec128} and
\ref{enc_dec16} show the encoding and decoding goodput in MB per
second for different generation sizes, $g = [1024, 128, 16]$
respectively. The size of each coded packet was fixed to $1536$~bytes
since that is the typical size of an Ethernet frame. The blocked operations
were performed dividing the matrices in squared sub-blocks of
$16,\ 32\ ,\ 64,\ldots,\ 1024$ operands (words in the Galois field) in
height and width. The figures show only block sizes of
$16 \times 16$, $32 \times 32$, $64 \times 64$ operands since with bigger
block sizes, the operands do not fit in L1 cache. Several testcases
are considered and detailed.

\subsubsection{Baseline encoding}
The baseline results involve no recording of the
\ac{DAG} and are performed in a \emph{by-the-book} fashion. The encoder uses
only one thread. The difference between the non-blocked and blocked
encoding schemes is that in the blocked scheme, the matrix multiplications are
performed dividing the matrices in sub-blocks in order to make the algorithm
cache efficient as described in Section~\ref{sub:implementation-multicore}.

\subsubsection{Encoding blocked}
The encoding results are obtained using the
method described in Section~\ref{sub:implementation-multicore}. The time
recorded includes the dependencies resolving, creation of the \ac{DAG},
and the taskz scheduling. In practice, it would suffice to calculate and
store this information only once per generation size.

@@

For $g= 1024$, the blocked baseline measurements outperforms the non blocked
variant. This means that making the matrix multiplication algorithm cache
efficient brings an increase in goodput by a factor of 2.83. When using the
algorithm described in Section~\ref{sub:implementation-multicore},
encoding with four cores is on average $3.7\times$ faster than with one core.
Similarly, decoding with four codes is $3.05\times$ faster, on average, than
decoding with a single core. Figure~\ref{enc_dec1024} shows that the
implemented algorithm, by exploiting cache efficiency and only three extra
cores provides a $10\times$ gain compared with traditional non-blocked
algorithms. With $g = 1024$, the matrix inversion becomes more expensive
than at smaller generations sizes. Therefore, the decoding goodput
is 46\% of the encoding goodput.

@@

For $g = 128$, the differences between the baselines operations show that a
blocked algorithm is 15\% faster than the non-blocked variant. Encoding with
four cores is $2.82\times$ faster than with a single core. Due to the smaller
matrix sizes, the gain when using blocked operations in the baselines is not
that significant when compared with $g = 1024$. For the same reason, the matrix
inversion is less expensive. As a consequence, the decoding goodput is 71\%
of the encoding goodput.

@@

The differences in goodput, for all generation sizes, between the
blocked baseline and the single threaded scheduled measurements are due the
time spent resolving the dependencies and the scheduling overhead. These
effects are negligible for big generation sizes, while considerable for
small matrices.

@@

To compare how much slower is the matrix multiplication with respect to the
matrix inversion for different generation sizes, we ran a set of tests. We used
a single core to perform the operations. We changed the generation sizes,
performed matrices multiplications and matrix inversions, and measured the time
spent doing so which we name $T_{mult}$ and $T_{inv}$. We calculate the ratio
between these two measured times defined as $r = \frac{T_{mult}}{T_{inv}}$.
Table~\ref{runtimes} summarizes the results. The bigger the matrix size, the
smaller is the calculated ratio. This means that when the problems are bigger,
the decoding goodput decreases compared with the encoding goodput.
